Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters. Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study. Our experiments show that while $\entropysgd$ trains significantly faster than SGD for recurrent networks, it gets relatively minor gains in terms of wall-clock time for CNNs. Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.

@@

This ``moving prior'' is absent in variational inference and indeed a crucial feature of the local entropy objective. The gradient of local entropy in Eqn.~\eqref{eq:entropysgd_grad} clarifies this point:
\aeqs{
    \nabla F(x, \g) &= -\g\ \rbrac{x - \ag{z;\ \Xi}} = -\g\ \E_{z\ \sim r(z; x)}\ \sqbrac{z};
}
where the distribution $r(z; x)$ is given by
$$
    r(z;\ x) \propto\ p(\Xi\ |\ z)\ \exp \rbrac{-\f{\g}{2}\ \norm{x-z}^2};
$$
it thus contains a data likelihood term along with a prior that ``moves'' along with the current iterate $x$.

Let us remark that methods in the deep learning literature that average the gradient in a local neighborhood of the weights $x$ through perturbations~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term) and are thus different from SVI and certainly, local entropy.

>>>>

If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters. Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study. Our experiments show that while $\entropysgd$ trains significantly faster than SGD for recurrent networks, it gets relatively minor gains in terms of wall-clock time for CNNs. Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate it in a teacher-student framework~\citep{balan2015bayesian} is another avenue for extensions to this work.

@@

This ``moving prior'' is absent in variational inference and indeed, a crucial feature of the local entropy objective. The gradient of local entropy in Eqn.~\eqref{eq:entropysgd_grad} clarifies this point:
\aeqs{
    \nabla F(x, \g) &= -\g\ \rbrac{x - \ag{z;\ \Xi}} = -\g\ \E_{z\ \sim r(z; x)}\ \sqbrac{z};
}
where the distribution $r(z; x)$ is given by
$$
    r(z;\ x) \propto\ p(\Xi\ |\ z)\ \exp \rbrac{-\f{\g}{2}\ \norm{x-z}^2};
$$
it thus contains a data likelihood term along with a prior that ``moves'' along with the current iterate $x$.

Let us remark that methods in the deep learning literature that average the gradient through perturbations in the neighborhood of $x$~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term); such an averaging it thus different from local entropy.
